---
title: "ICE-A: Figure 1"
output: 
  html_notebook
---
# 1 Proximity vs interaction based annotation

## 1.0 Preperations
Run before:
conda activate ice_analysis_env
export LD_LIBRARY_PATH=$LD_LIBRARY_PATH:/opt/anaconda3/envs/icea_analysis_env/lib

Setting working directory
```{r setup}
knitr::opts_knit$set(root.dir = normalizePath("/mnt/data/common/tobias/tg/ICE_A/ICEA_analysis/"))
```


Loading packages
```{r}
packages <- c("AnnotationDbi","org.Mm.eg.db", "clusterProfiler", "ggplot2", "ggpubr", "readxl", "tidyr", "tilingArray", "dplyr", "mltools", "stringr")
lapply(packages, library, character.only = TRUE)
```

## 1.1 Intro to ICEA
Explanatory figure of proximity vs interaction-based annotation

## 1.2 Limitation of proximity

### 1.2.1 Preproceesing for calcualtion of proximity/GREAT distance limitations
TSS positions from: https://great-help.atlassian.net/wiki/spaces/GREAT/pages/655445/Genes
Chom sizes from UCSC, filtered for chr 1-19, x,y
Mouse (mm10)
```{r}
#Load tss regions from great
tss.mm10 <- read.table("data/other/tss/GREATv4.genes.mm10.tsv", header=F, sep="\t")
colnames(tss.mm10) <- c("ens", "chr", "tss", "strand", "gene")
tss.mm10$type <- "gene"
tss.mm10 <- tss.mm10[tss.mm10$chr != "chrM",]

#Adding start/end positions
chr.size.mm10 <- read.table("data/other/chr_sizes/mm10.chrom.sizes", header=F, sep="\t")
colnames(chr.size.mm10) <- c("Chr", "Length")
chr.start.end.mm10 <- rbind(data.frame(ens=rep("", nrow(chr.size.mm10)), chr= chr.size.mm10$Chr, tss=0, strand=rep("_",nrow(chr.size.mm10)), gene=rep("", nrow(chr.size.mm10)), type="start"), data.frame(ens=rep("", nrow(chr.size.mm10)), chr= chr.size.mm10$Chr, tss=chr.size.mm10$Length, strand=rep("_",nrow(chr.size.mm10)), gene=rep("", nrow(chr.size.mm10)), type="end"))
tss.mm10 <- rbind(tss.mm10,chr.start.end.mm10 )

#Proximity: Calculate gene regulatory domains
tss.mm10.prox <- tss.mm10
tss.mm10.prox <- tss.mm10.prox[order(tss.mm10.prox$chr, tss.mm10.prox$tss),]
for (r in 1:nrow(tss.mm10)){
  tss.mm10.prox[r, "dist_up"] <- ifelse(tss.mm10.prox[r,"strand"]=="+", ifelse(tss.mm10.prox[r-1,"type"]=="gene", (tss.mm10.prox[r,"tss"]-tss.mm10.prox[r-1,"tss"])/2, (tss.mm10.prox[r,"tss"]-tss.mm10.prox[r-1,"tss"])), ifelse(tss.mm10.prox[r+1,"type"]=="gene", -(tss.mm10.prox[r,"tss"]-tss.mm10.prox[r+1,"tss"])/2, -(tss.mm10.prox[r,"tss"]-tss.mm10.prox[r+1,"tss"])))
  tss.mm10.prox[r, "dist_down"] <- ifelse(tss.mm10.prox[r,"strand"]=="+", ifelse(tss.mm10.prox[r+1,"type"]=="gene", -(tss.mm10.prox[r,"tss"]-tss.mm10.prox[r+1,"tss"])/2, -(tss.mm10.prox[r,"tss"]-tss.mm10.prox[r+1,"tss"])), ifelse(tss.mm10.prox[r-1,"type"]=="gene", (tss.mm10.prox[r,"tss"]-tss.mm10.prox[r-1,"tss"])/2, (tss.mm10.prox[r,"tss"]-tss.mm10.prox[r-1,"tss"])))
tss.mm10.prox[r, "min"] <- ifelse(tss.mm10.prox[r,"strand"]=="+", tss.mm10.prox[r,]$tss-tss.mm10.prox[r,]$dist_up, tss.mm10.prox[r,]$tss-tss.mm10.prox[r,]$dist_down)
tss.mm10.prox[r, "max"] <- ifelse(tss.mm10.prox[r,"strand"]=="+", tss.mm10.prox[r,]$tss+tss.mm10.prox[r,]$dist_down, tss.mm10.prox[r,]$tss+tss.mm10.prox[r,]$dist_up)
}


#GREAT: Calculate gene regulatory domains
tss.mm10.great <- tss.mm10
tss.mm10.great$tss_start <- ifelse(tss.mm10.great$strand=="+", tss.mm10.great$tss-5000, tss.mm10.great$tss-1000)
tss.mm10.great$tss_start <- ifelse(tss.mm10.great$strand=="_", tss.mm10.great$tss, tss.mm10.great$tss_start)
tss.mm10.great$tss_end <- ifelse(tss.mm10.great$strand=="+", tss.mm10.great$tss+1000, tss.mm10.great$tss+5000)
tss.mm10.great$tss_end <- ifelse(tss.mm10.great$strand=="_", tss.mm10.great$tss, tss.mm10.great$tss_end)
tss.mm10.great <- tss.mm10.great[order(tss.mm10.great$chr, tss.mm10.great$tss),]

for (r in 1:nrow(tss.mm10.great)){
tss.mm10.great[r, "dist_up"] <- ifelse(tss.mm10.great[r,"strand"]=="+", (tss.mm10.great[r,"tss"]-tss.mm10.great[r-1,"tss_end"]), -(tss.mm10.great[r,"tss"]-tss.mm10.great[r+1,"tss_start"]))
  tss.mm10.great[r, "dist_up"] <- ifelse(tss.mm10.great[r, "dist_up"]<5000, 5000, tss.mm10.great[r, "dist_up"])

tss.mm10.great[r, "dist_down"] <- ifelse(tss.mm10.great[r,"strand"]=="+",  -(tss.mm10.great[r,"tss"]-tss.mm10.great[r+1,"tss_start"]), (tss.mm10.great[r,"tss"]-tss.mm10.great[r-1,"tss_end"]))
tss.mm10.great[r, "dist_down"] <- ifelse(tss.mm10.great[r, "dist_down"]<1000, 1000, tss.mm10.great[r, "dist_down"])
    
tss.mm10.great[r, "min"] <- ifelse(tss.mm10.great[r,"strand"]=="+", tss.mm10.great[r,]$tss-tss.mm10.great[r,]$dist_up, tss.mm10.great[r,]$tss-tss.mm10.great[r,]$dist_down)
tss.mm10.great[r, "max"] <- ifelse(tss.mm10.great[r,"strand"]=="+", tss.mm10.great[r,]$tss+tss.mm10.great[r,]$dist_down, tss.mm10.great[r,]$tss+tss.mm10.great[r,]$dist_up)
}

#Add curated regions
great.curated.regions<- read.table("data/other/tss/GREATv4.curatedRegDoms.mm10.txt", header=F, sep="\t")
colnames(great.curated.regions) <- c("chr", "start","end", "ens", "entrez")
for (i in great.curated.regions$ens){
  tss.mm10.great[tss.mm10.great$ens %in% i, "min"] <- ifelse(tss.mm10.great[tss.mm10.great$ens %in% i,"min"]<great.curated.regions[great.curated.regions$ens %in% i,"start"], tss.mm10.great[tss.mm10.great$ens %in% i,"min"], great.curated.regions[great.curated.regions$ens %in% i,"start"])
  tss.mm10.great[tss.mm10.great$ens %in% i, "max"] <- ifelse(tss.mm10.great[tss.mm10.great$ens %in% i,"max"]>great.curated.regions[great.curated.regions$ens %in% i,"end"], tss.mm10.great[tss.mm10.great$ens %in% i,"max"], great.curated.regions[great.curated.regions$ens %in% i,"end"])
  tss.mm10.great[tss.mm10.great$ens %in% i, "dist_up"]  <- ifelse(tss.mm10.great[tss.mm10.great$ens %in% i, "strand"]=="+", (tss.mm10.great[tss.mm10.great$ens %in% i, "tss"]-tss.mm10.great[tss.mm10.great$ens %in% i, "min"]), (tss.mm10.great[tss.mm10.great$ens %in% i, "max"]-tss.mm10.great[tss.mm10.great$ens %in% i, "tss"]))
  tss.mm10.great[tss.mm10.great$ens %in% i, "dist_down"]  <- ifelse(tss.mm10.great[tss.mm10.great$ens %in% i, "strand"]=="+", (tss.mm10.great[tss.mm10.great$ens %in% i, "max"]-tss.mm10.great[tss.mm10.great$ens %in% i, "tss"]), (tss.mm10.great[tss.mm10.great$ens %in% i, "tss"]-tss.mm10.great[tss.mm10.great$ens %in% i, "min"]))
}
```



Human (Hg38)
```{r}
#Load tss regions from great
tss.hg38 <- read.table("data/other/tss/GREATv4.genes.hg38.tsv", header=F, sep="\t")
colnames(tss.hg38) <- c("ens", "chr", "tss", "strand", "gene")
tss.hg38$type <- "gene"
tss.hg38 <- tss.hg38[tss.hg38$chr != "chrM",]

#Adding start/end positions
chr.size.hg38 <- read.table("data/other/chr_sizes/hg38.chrom.sizes", header=F, sep="\t")
colnames(chr.size.hg38) <- c("Chr", "Length")
chr.start.end.hg38 <- rbind(data.frame(ens=rep("", nrow(chr.size.hg38)), chr= chr.size.hg38$Chr, tss=0, strand=rep("_",nrow(chr.size.hg38)), gene=rep("", nrow(chr.size.hg38)), type="start"), data.frame(ens=rep("", nrow(chr.size.hg38)), chr= chr.size.hg38$Chr, tss=chr.size.hg38$Length, strand=rep("_",nrow(chr.size.hg38)), gene=rep("", nrow(chr.size.hg38)), type="end"))
tss.hg38 <- rbind(tss.hg38,chr.start.end.hg38 )

#Proximity: Calculate gene regulatory domains
tss.hg38.prox <- tss.hg38
tss.hg38.prox <- tss.hg38.prox[order(tss.hg38.prox$chr, tss.hg38.prox$tss),]
for (r in 1:nrow(tss.hg38)){
  tss.hg38.prox[r, "dist_up"] <- ifelse(tss.hg38.prox[r,"strand"]=="+", ifelse(tss.hg38.prox[r-1,"type"]=="gene", (tss.hg38.prox[r,"tss"]-tss.hg38.prox[r-1,"tss"])/2, (tss.hg38.prox[r,"tss"]-tss.hg38.prox[r-1,"tss"])), ifelse(tss.hg38.prox[r+1,"type"]=="gene", -(tss.hg38.prox[r,"tss"]-tss.hg38.prox[r+1,"tss"])/2, -(tss.hg38.prox[r,"tss"]-tss.hg38.prox[r+1,"tss"])))
  tss.hg38.prox[r, "dist_down"] <- ifelse(tss.hg38.prox[r,"strand"]=="+", ifelse(tss.hg38.prox[r+1,"type"]=="gene", -(tss.hg38.prox[r,"tss"]-tss.hg38.prox[r+1,"tss"])/2, -(tss.hg38.prox[r,"tss"]-tss.hg38.prox[r+1,"tss"])), ifelse(tss.hg38.prox[r-1,"type"]=="gene", (tss.hg38.prox[r,"tss"]-tss.hg38.prox[r-1,"tss"])/2, (tss.hg38.prox[r,"tss"]-tss.hg38.prox[r-1,"tss"])))
tss.hg38.prox[r, "min"] <- ifelse(tss.hg38.prox[r,"strand"]=="+", tss.hg38.prox[r,]$tss-tss.hg38.prox[r,]$dist_up, tss.hg38.prox[r,]$tss-tss.hg38.prox[r,]$dist_down)
tss.hg38.prox[r, "max"] <- ifelse(tss.hg38.prox[r,"strand"]=="+", tss.hg38.prox[r,]$tss+tss.hg38.prox[r,]$dist_down, tss.hg38.prox[r,]$tss+tss.hg38.prox[r,]$dist_up)
}

#GREAT: Calculate gene regulatory domains
tss.hg38.great <- tss.hg38
tss.hg38.great$tss_start <- ifelse(tss.hg38.great$strand=="+", tss.hg38.great$tss-5000, tss.hg38.great$tss-1000)
tss.hg38.great$tss_start <- ifelse(tss.hg38.great$strand=="_", tss.hg38.great$tss, tss.hg38.great$tss_start)
tss.hg38.great$tss_end <- ifelse(tss.hg38.great$strand=="+", tss.hg38.great$tss+1000, tss.hg38.great$tss+5000)
tss.hg38.great$tss_end <- ifelse(tss.hg38.great$strand=="_", tss.hg38.great$tss, tss.hg38.great$tss_end)
tss.hg38.great <- tss.hg38.great[order(tss.hg38.great$chr, tss.hg38.great$tss),]

for (r in 1:nrow(tss.hg38.great)){
tss.hg38.great[r, "dist_up"] <- ifelse(tss.hg38.great[r,"strand"]=="+", (tss.hg38.great[r,"tss"]-tss.hg38.great[r-1,"tss_end"]), -(tss.hg38.great[r,"tss"]-tss.hg38.great[r+1,"tss_start"]))
  tss.hg38.great[r, "dist_up"] <- ifelse(tss.hg38.great[r, "dist_up"]<5000, 5000, tss.hg38.great[r, "dist_up"])

tss.hg38.great[r, "dist_down"] <- ifelse(tss.hg38.great[r,"strand"]=="+",  -(tss.hg38.great[r,"tss"]-tss.hg38.great[r+1,"tss_start"]), (tss.hg38.great[r,"tss"]-tss.hg38.great[r-1,"tss_end"]))
tss.hg38.great[r, "dist_down"] <- ifelse(tss.hg38.great[r, "dist_down"]<1000, 1000, tss.hg38.great[r, "dist_down"])
    
tss.hg38.great[r, "min"] <- ifelse(tss.hg38.great[r,"strand"]=="+", tss.hg38.great[r,]$tss-tss.hg38.great[r,]$dist_up, tss.hg38.great[r,]$tss-tss.hg38.great[r,]$dist_down)
tss.hg38.great[r, "max"] <- ifelse(tss.hg38.great[r,"strand"]=="+", tss.hg38.great[r,]$tss+tss.hg38.great[r,]$dist_down, tss.hg38.great[r,]$tss+tss.hg38.great[r,]$dist_up)
}

#Add curated regions
great.curated.regions<- read.table("data/other/tss/GREATv4.curatedRegDoms.hg38.txt", header=F, sep="\t")
colnames(great.curated.regions) <- c("chr", "start","end", "ens", "entrez")
for (i in great.curated.regions$ens){
  tss.hg38.great[tss.hg38.great$ens %in% i, "min"] <- ifelse(tss.hg38.great[tss.hg38.great$ens %in% i,"min"]<great.curated.regions[great.curated.regions$ens %in% i,"start"], tss.hg38.great[tss.hg38.great$ens %in% i,"min"], great.curated.regions[great.curated.regions$ens %in% i,"start"])
  tss.hg38.great[tss.hg38.great$ens %in% i, "max"] <- ifelse(tss.hg38.great[tss.hg38.great$ens %in% i,"max"]>great.curated.regions[great.curated.regions$ens %in% i,"end"], tss.hg38.great[tss.hg38.great$ens %in% i,"max"], great.curated.regions[great.curated.regions$ens %in% i,"end"])
  tss.hg38.great[tss.hg38.great$ens %in% i, "dist_up"]  <- ifelse(tss.hg38.great[tss.hg38.great$ens %in% i, "strand"]=="+", (tss.hg38.great[tss.hg38.great$ens %in% i, "tss"]-tss.hg38.great[tss.hg38.great$ens %in% i, "min"]), (tss.hg38.great[tss.hg38.great$ens %in% i, "max"]-tss.hg38.great[tss.hg38.great$ens %in% i, "tss"]))
  tss.hg38.great[tss.hg38.great$ens %in% i, "dist_down"]  <- ifelse(tss.hg38.great[tss.hg38.great$ens %in% i, "strand"]=="+", (tss.hg38.great[tss.hg38.great$ens %in% i, "max"]-tss.hg38.great[tss.hg38.great$ens %in% i, "tss"]), (tss.hg38.great[tss.hg38.great$ens %in% i, "tss"]-tss.hg38.great[tss.hg38.great$ens %in% i, "min"]))
}
```

Calculate statistics
```{r}
#Calculate statistics
#Mouse
tss.mm10.prox <- tss.mm10.prox[tss.mm10.prox$type=="gene",]
tss.mm10.prox$reg <- tss.mm10.prox$dist_up+tss.mm10.prox$dist_down
tss.mm10.prox$maxdist <- pmax(tss.mm10.prox$dist_up, tss.mm10.prox$dist_down)
tss.mm10.prox$mindist <- pmin(tss.mm10.prox$dist_up, tss.mm10.prox$dist_down)
tss.mm10.prox$meandist <- rowMeans(tss.mm10.prox[,c("mindist", "maxdist")])
tss.mm10.prox$anno="PROXIMITY"
tss.mm10.prox$species="Mouse"


tss.mm10.great <- tss.mm10.great[tss.mm10.great$type=="gene",]
tss.mm10.great$reg <- tss.mm10.great$dist_up+tss.mm10.great$dist_down
tss.mm10.great$maxdist <- pmax(tss.mm10.great$dist_up, tss.mm10.great$dist_down)
tss.mm10.great$mindist <- pmin(tss.mm10.great$dist_up, tss.mm10.great$dist_down)
tss.mm10.great$meandist <- rowMeans(tss.mm10.great[,c("mindist", "maxdist")])
tss.mm10.great$anno="GREAT"
tss.mm10.great$species="Mouse"

#Huamn
tss.hg38.prox <- tss.hg38.prox[tss.hg38.prox$type=="gene",]
tss.hg38.prox$reg <- tss.hg38.prox$dist_up+tss.hg38.prox$dist_down
tss.hg38.prox$maxdist <- pmax(tss.hg38.prox$dist_up, tss.hg38.prox$dist_down)
tss.hg38.prox$mindist <- pmin(tss.hg38.prox$dist_up, tss.hg38.prox$dist_down)
tss.hg38.prox$meandist <- rowMeans(tss.hg38.prox[,c("mindist", "maxdist")])
tss.hg38.prox$anno="PROXIMITY"
tss.hg38.prox$species="Human"

tss.hg38.great <- tss.hg38.great[tss.hg38.great$type=="gene",]
tss.hg38.great$reg <- tss.hg38.great$dist_up+tss.hg38.great$dist_down
tss.hg38.great$maxdist <- pmax(tss.hg38.great$dist_up, tss.hg38.great$dist_down)
tss.hg38.great$mindist <- pmin(tss.hg38.great$dist_up, tss.hg38.great$dist_down)
tss.hg38.great$meandist <- rowMeans(tss.hg38.great[,c("mindist", "maxdist")])
tss.hg38.great$anno="GREAT"
tss.hg38.great$species="Human"


#Print statistics
#Mouse
print("Mouse (Mm10)")
print(paste0("Median max distance Proximity: ",median(tss.mm10.prox[!is.na(tss.mm10.prox$maxdist),]$maxdist)))
print(paste0("Median max distance GREAT: ",median(tss.mm10.great[!is.na(tss.mm10.great$maxdist),]$maxdist)))

print(paste0("Median min distance Proximity: ",median(tss.mm10.prox[!is.na(tss.mm10.prox$maxdist),]$mindist)))
print(paste0("Median min distance GREAT: ",median(tss.mm10.great[!is.na(tss.mm10.great$maxdist),]$mindist)))

print(paste0("Median average distance Proximity: ",median(tss.mm10.prox[!is.na(tss.mm10.prox$maxdist),]$meandist)))
print(paste0("Median average distance GREAT: ",median(tss.mm10.great[!is.na(tss.mm10.great$maxdist),]$meandist)))

print(paste0("Median upstream distance Proximity: ",median(tss.mm10.prox[!is.na(tss.mm10.prox$maxdist),]$dist_up)))
print(paste0("Median upstream distance GREAT: ",median(tss.mm10.great[!is.na(tss.mm10.great$maxdist),]$dist_up)))

print(paste0("Median downstream distance Proximity: ",median(tss.mm10.prox[!is.na(tss.mm10.prox$maxdist),]$dist_down)))
print(paste0("Median downstream distance GREAT: ",median(tss.mm10.great[!is.na(tss.mm10.great$maxdist),]$dist_down)))

print(paste0("% over 100kb Proximity: ",nrow(tss.mm10.prox[!is.na(tss.mm10.prox$maxdist) & tss.mm10.prox$maxdist>100000,])/nrow(tss.mm10.prox[!is.na(tss.mm10.prox$maxdist),])))
print(paste0("% over 100kb GREAT: ",nrow(tss.mm10.great[!is.na(tss.mm10.great$maxdist) & tss.mm10.great$maxdist>100000,])/nrow(tss.mm10.great[!is.na(tss.mm10.great$maxdist),])))

#Human
print("Human (Hg38)")
print(paste0("Median max distance Proximity: ",median(tss.hg38.prox[!is.na(tss.hg38.prox$maxdist),]$maxdist)))
print(paste0("Median max distance GREAT: ",median(tss.hg38.great[!is.na(tss.hg38.great$maxdist),]$maxdist)))

print(paste0("Median min distance Proximity: ",median(tss.hg38.prox[!is.na(tss.hg38.prox$maxdist),]$mindist)))
print(paste0("Median min distance GREAT: ",median(tss.hg38.great[!is.na(tss.hg38.great$maxdist),]$mindist)))

print(paste0("Median average distance Proximity: ",median(tss.hg38.prox[!is.na(tss.hg38.prox$maxdist),]$meandist)))
print(paste0("Median average distance GREAT: ",median(tss.hg38.great[!is.na(tss.hg38.great$maxdist),]$meandist)))

print(paste0("Median upstream distance Proximity: ",median(tss.hg38.prox[!is.na(tss.hg38.prox$maxdist),]$dist_up)))
print(paste0("Median upstream distance GREAT: ",median(tss.hg38.great[!is.na(tss.hg38.great$maxdist),]$dist_up)))

print(paste0("Median downstream distance Proximity: ",median(tss.hg38.prox[!is.na(tss.hg38.prox$maxdist),]$dist_down)))
print(paste0("Median downstream distance GREAT: ",median(tss.hg38.great[!is.na(tss.hg38.great$maxdist),]$dist_down)))

print(paste0("% over 100kb Proximity: ",nrow(tss.hg38.prox[!is.na(tss.hg38.prox$maxdist) & tss.hg38.prox$maxdist>100000,])/nrow(tss.hg38.prox[!is.na(tss.hg38.prox$maxdist),])))
print(paste0("% over 100kb GREAT: ",nrow(tss.hg38.great[!is.na(tss.hg38.great$maxdist) & tss.hg38.great$maxdist>100000,])/nrow(tss.hg38.great[!is.na(tss.hg38.great$maxdist),])))

```

Merge data
```{r}
tss.comb <- rbind(tss.mm10.prox, tss.mm10.great[,c(1:6, 9:18)], tss.hg38.prox, tss.hg38.great[,c(1:6, 9:18)])
tss.comb$anno <- factor(tss.comb$anno, levels=c("PROXIMITY", "GREAT"))
tss.comb$species <- factor(tss.comb$species, levels=c("Mouse", "Human"))
tss.comb[, c(7:8,11:14)][tss.comb[, c(7:8,11:14)] == 0] <- 1
```


### 1.2.2 Max distance for CRE annotation
Violin plot of max  distance (Figure 1B)
```{r}
plot.violin.maxdistance<-ggplot(tss.comb, aes(x=anno, y=dist_down, fill=anno))+
  geom_violin(color="black")+
  annotate("rect", xmin=0, xmax=3, ymin=0, ymax=100000, alpha=0.6, fill="white")+
  geom_violin(color="black", alpha=0)+
  geom_boxplot(width=0.1, color="black", alpha=0)+
  geom_hline(yintercept=100000, linetype="dashed" ) + 
  scale_y_log10(breaks=c(10, 100, 1000, 10000, 100000, 1000000), labels= c("0.01 kb ", "0.1 kb ", "1 kb ", "10 kb ", "100 kb ", "1 Mb "), limits=c(1,1e7), expand=c(0,0))+TG_theme()+
  xlab("")+ylab("Maximum distance for\nCRE annotation")+
  theme(axis.title.y=element_text(size=20), axis.text.x = element_text(size=14, colour ="black", angle = 45, hjust = 1, vjust = 0.9), axis.text.y = element_text(size=14, colour ="black"), legend.position = "none", aspect.ratio = 1.75)+
    scale_fill_manual(name="Annotation",values = c("grey75", "grey25"))+
  facet_grid(. ~ species)


ggsave(path="output/plots/figure1", filename="Max_distance_threshold_for_CRE_annotation.pdf", plot = plot.violin.maxdistance, device = "pdf", dpi=300)
```


### 1.2.3 Median upper distance for CRE annotation
Barplot of median upper distance (Figure 1C)
```{r}
tss.comb.up.down <- rbind(data.frame(tss.comb[,c(1:6,15,16)], dist=tss.comb$dist_up, direction="Upstream"), data.frame(tss.comb[,c(1:6,15,16)], dist=tss.comb$dist_down, direction="Downstream"))
tss.comb.up.down$direction <- factor(tss.comb.up.down$direction, levels=c("Upstream", "Downstream"))
tss.comb.up.down.median <- aggregate(dist ~ anno+ species+ direction, data=tss.comb.up.down[,7:10], FUN=median)

source("script/general/r_functions/TG_theme.R")
plot.box.updowndist<-ggplot(tss.comb.up.down.median , aes(x=anno, y=dist, fill=anno))+
  geom_bar(stat="identity", color="black", width = 0.75)+
  scale_y_continuous(breaks=c(0,25000, 50000), labels= c("0 kb ", "25 kb ", "50 kb "), limits=c(0,65000), expand=c(0,0))+
  xlab("")+ylab("Maximum distance for\nCRE annotation")+
  TG_theme(aspect.ratio = 4)+
  theme(axis.title.y=element_text(size=20), axis.text.x = element_text(size=14, colour ="black", angle = 45, hjust = 1, vjust = 0.9), axis.text.y = element_text(size=14, colour ="black"), legend.position = "none", aspect.ratio = 1.75)+
    scale_fill_manual(name="Annotation",values = c("grey75", "grey25"))+
  facet_grid(species ~ direction)

ggsave(path="output/plots/figure1", filename="Distance_threshold_for_CRE_annotation.pdf", plot = plot.box.updowndist, device = "pdf", dpi=300)
```


## 1.3 Benchmarking using validated enhancers in K562 cells

### 1.3.1 Indetification of experimentally validadted enhancers
Load and filter significant enhancers from Fulco et al. 2019
```{r}
# Load Enhancer.Promoter interaction 
Enh.Prom.pairs <- read_excel("data/other/K562_enhancers/41588_2019_538_MOESM3_ESM.xlsx", sheet="Supplementary Table 3a", skip=1)
Enh.Prom.pairs$peakID <- paste(Enh.Prom.pairs$chr, ":", Enh.Prom.pairs$start, "-", Enh.Prom.pairs$end, sep = "")

# Filter valid enhancer-gene pairs (Enhancers overlaping bounds of gene removed)
Enh.Prom.pairs <- Enh.Prom.pairs[Enh.Prom.pairs$`Valid E-G connection` == TRUE,]

# Loading gene info and merging with enhancer-gene pairs
Gene.info <- read_excel("data/other/K562_enhancers/41588_2019_538_MOESM3_ESM.xlsx", sheet="Supplementary Table 5b", skip=1)
Enh.Prom.pairs <- merge(Enh.Prom.pairs, Gene.info, by="Gene", all.x=TRUE, suffixes=c("", "_TSS"))

# Calculate distance between peak center and corresponding TSS 
Enh.Prom.pairs$TSS <- as.numeric(ifelse(Enh.Prom.pairs$strand == "+", Enh.Prom.pairs$start_TSS, Enh.Prom.pairs$end_TSS))
Enh.Prom.pairs$DistanceTSS <- ((Enh.Prom.pairs$start + Enh.Prom.pairs$end)/2)-Enh.Prom.pairs$TSS 

# Filter pairs based on distance to TSS: Max distance to TSS set to 3Mb (based on FitHiChIP analysis)
Enh.Prom.pairs <- Enh.Prom.pairs[abs(Enh.Prom.pairs$DistanceTSS) <= 3000000,]
Enh.Prom.pairs$id <- paste0(Enh.Prom.pairs$peakID, "_", Enh.Prom.pairs$Gene)
Enh.Prom.pairs <- Enh.Prom.pairs[!duplicated(Enh.Prom.pairs$id),]

#Distance bins
Enh.Prom.pairs$DistanceBin <- bin_data(abs(Enh.Prom.pairs$DistanceTSS), bins=c(0, 10000, 50000, 150000,3000000), binType = "explicit")

# Writing bed files for filtered enhancer-gene pairs for annotation
Enh.Prom.pairs.bed <- unique(Enh.Prom.pairs[,c("chr", "start", "end", "peakID")])
write.table(Enh.Prom.pairs.bed, file="output/data/figure1/K562_Enhancers_filtered.bed", quote=FALSE, sep="\t", col.names=FALSE, row.names=FALSE)

# Significant enhancer-gene pairs
Enh.Prom.pairs.sign <- Enh.Prom.pairs[Enh.Prom.pairs$Significant==TRUE,]
```


### 1.3.2 Annotation of OCRs
HOMER proximity annotation 
```{bash engine.opts='-l'}
cd output/data/figure1/annotation
annotatePeaks.pl ../K562_Enhancers_filtered.bed hg19 > K562_Enhancers_HOMER_annotated.txt
awk -v FS='\t' -v OFS='\t' '{print $1, $16, $10}' K562_Enhancers_HOMER_annotated.txt > K562_Enhancers_HOMER_annotated_cut.txt
```

```{r}
Enh.anno.HOMER <- read.table("output/data/figure1/annotation/K562_Enhancers_HOMER_annotated_cut.txt", header=TRUE, sep="\t", quote="")
names(Enh.anno.HOMER)[1] <- 'peakID'
Enh.anno.HOMER$id <- paste0(Enh.anno.HOMER$peakID, "_", Enh.anno.HOMER$Gene)
Enh.anno.HOMER <- Enh.anno.HOMER[!duplicated(Enh.anno.HOMER$id),]
Enh.anno.HOMER <- Enh.anno.HOMER[Enh.anno.HOMER$Gene.Name !="",]
```

GREAT annotation
Run great using default settings (except Default settings
```{r}
Enh.anno.GREAT <- read.table("output/data/figure1/annotation/K562_Enhancers_GREAT_annotated.txt", header=F, sep="\t", quote="")
colnames(Enh.anno.GREAT) <- c("peakID", "Gene")
Enh.anno.GREAT <- mutate(Enh.anno.GREAT,Gene = strsplit(Gene, ", ")) %>% unnest(Gene) %>% separate(Gene, c('Gene', 'DistanceTSS'), sep=" \\(")
Enh.anno.GREAT$DistanceTSS <- str_replace(Enh.anno.GREAT$DistanceTSS,
                        pattern = "\\)",
                        replacement = "")
Enh.anno.GREAT$DistanceTSS <- as.numeric(Enh.anno.GREAT$DistanceTSS)
Enh.anno.GREAT$id <- paste0(Enh.anno.GREAT$peakID, "_", Enh.anno.GREAT$Gene)
Enh.anno.GREAT <- Enh.anno.GREAT[!duplicated(Enh.anno.GREAT$id),]
```


ICE-A interaction-based annotation of OCRs
```{r}
int <- read.table("data/interactions/GSE101498_K562_HiChIP_H3K27ac_FitHiChIP_v_9_1.interactions_FitHiC_Q0.05_MergeNearContacts.bed", header=TRUE, sep="\t", quote="")

int$score <- -log(int$Q.Value_Bias)
int[is.infinite(int$score),"score"] <- 1000
int <- int[order(int$score, decreasing = T),]
int.score.50 <- min(int[1:nrow(int)*0.5,]$score)
int.score.20 <- min(int[1:nrow(int)*0.20,]$score)
int.score.5 <- min(int[1:nrow(int)*0.05,]$score)
```

```{bash engine.opts='-l'}
script_dir="script/figure1"
bash $script_dir/ICEA_GSE101498_K562_HiChIP_H3K27ac_K562_enhancers_annotation.sh
```

```{r}
Enh.anno.ICEA <- read.table("output/data/figure1/annotation/K562_enhancers_HiChIP_H3K27ac_annotation/Peak_annotation/K562_enhancers/K562_enhancers_PLACseq_annotated.txt", header=TRUE, sep="\t", quote="")

Enh.anno.ICEA$peakID <- paste(Enh.anno.ICEA$Chr, ":", Enh.anno.ICEA$Start, "-", Enh.anno.ICEA$End, sep = "")
Enh.anno.ICEA$id <- paste0(Enh.anno.ICEA$peakID, "_", Enh.anno.ICEA$Gene)
Enh.anno.ICEA <- Enh.anno.ICEA[!duplicated(Enh.anno.ICEA$id),]
Enh.anno.ICEA <- Enh.anno.ICEA[!is.na(Enh.anno.ICEA$Distance_to_TSS),]

```

Overlap only
```{r}
Enh.anno.noprox <- Enh.anno.ICEA[Enh.anno.ICEA$Annotation_method=="Interaction_anno",]
Enh.anno.noneaby <- Enh.anno.ICEA[!(Enh.anno.ICEA$Annotation_method=="Interaction_anno" & abs(Enh.anno.ICEA$Peak_bin)>0) | !(Enh.anno.ICEA$Annotation_method=="Interaction_anno" & abs(Enh.anno.ICEA$Promoter_bin)>0),]
Enh.anno.noneaby <- Enh.anno.ICEA[(is.na(Enh.anno.ICEA$Peak_bin) | Enh.anno.ICEA$Peak_bin==0) & (is.na(Enh.anno.ICEA$Promoter_bin) | Enh.anno.ICEA$Promoter_bin==0),]
Enh.anno.overlap <- Enh.anno.noneaby[Enh.anno.noneaby$Annotation_method=="Interaction_anno",]

```


### 1.3.3 Comparison in the ability to indetify enhancers in distance ranges
Calculate percentage of identified enhancers, split into distance ranges
```{r}
Enh.Prom.pairs.sig <- Enh.Prom.pairs[Enh.Prom.pairs$Significant==1,]
Enh.Prom.pairs.sig$PROXIMITY <- ifelse(Enh.Prom.pairs.sig$id %in% Enh.anno.HOMER$id, 1, 0)
Enh.Prom.pairs.sig$GREAT <- ifelse(Enh.Prom.pairs.sig$id %in% Enh.anno.GREAT$id, 1, 0)
Enh.Prom.pairs.sig$ICEA <- ifelse(Enh.Prom.pairs.sig$id %in% Enh.anno.ICEA$id, 1, 0)
Enh.Prom.pairs.sig$NoProx <- ifelse(Enh.Prom.pairs.sig$id %in% Enh.anno.noprox$id, 1, 0)
#Enh.Prom.pairs.sig$NoNearby <- ifelse(Enh.Prom.pairs.sig$id %in% Enh.anno.noneaby$id, 1, 0)
Enh.Prom.pairs.sig$Overlap <- ifelse(Enh.Prom.pairs.sig$id %in% Enh.anno.overlap$id, 1, 0)

Enh.Prom.pairs.sig.summary <- gather(Enh.Prom.pairs.sig[c(1,27:33)], method, count, PROXIMITY:Overlap, factor_key=TRUE)
Enh.Prom.pairs.sig.summary$method <- factor(Enh.Prom.pairs.sig.summary$method, levels=c("PROXIMITY", "GREAT","ICEA", "NoProx","Overlap"))
Enh.Prom.pairs.sig.summary.element <- Enh.Prom.pairs.sig.summary %>% 
  dplyr::group_by(DistanceBin, method) %>% 
  dplyr::count(count, .drop=F) %>% 
  dplyr::mutate(percent= n/sum(n))
```

Barplot
```{r}
dist.lab <- c("<10kb", "10-50kb", "50-150kb", ">150kp") 
names(dist.lab) = levels(Enh.Prom.pairs.sig.summary$DistanceBin)

Enh.plot <-ggplot(data=Enh.Prom.pairs.sig.summary.element[Enh.Prom.pairs.sig.summary.element$count==1,], aes(x=method, y=percent, fill=method)) +
  geom_bar(stat="identity", position=position_dodge(width = 0.5), width=0.8, color="black")+
  facet_wrap(. ~ DistanceBin, nrow=1,   labeller = labeller(DistanceBin = dist.lab), strip.position = "bottom")+
  scale_x_discrete(breaks=c("PROXIMITY", "GREAT","ICEA", "Overlap"), labels=c("", "", "", ""))+
  xlab("")+
  scale_y_continuous(labels=scales::percent_format(accuracy = 1), name="% Enhancers identified", expand = c(0,0))+
  labs(y = "Percent", fill="method")+
  TG_theme(aspect.ratio=3)+
  scale_fill_viridis_d(name="Annotation",alpha = 1,begin = 0, end = 1,direction = -1,option = "D",aesthetics = "fill")+
  theme(axis.title.y=element_text(size=20), axis.text.y = element_text(size=14), legend.position="right", legend.title = element_blank(), legend.text = element_text(size = 10),  strip.text.x =element_text(size = 16),
        strip.placement = "outside",  legend.box.margin=margin(-200,0,-25,-100))+
  scale_fill_manual(values = c("#D1D1D3", "black", "#78A056", "orange"))

ggsave(path="output/plots/figure1", filename="K562_enhancer_annotation_element_based.pdf", plot = Enh.plot, device = "pdf", dpi=300)

```

### 1.3.4 Gene level enhancer indetificantion increased with “ICEA”
Calculate  enhancer identification per gene
```{r}
Enh.Prom.pairs.sig.summary.gene.prom <- Enh.Prom.pairs.sig.summary[Enh.Prom.pairs.sig.summary$DistanceBin=="[0, 10000)",] %>% 
  dplyr::group_by(Gene, method) %>% 
  dplyr::count(count, .drop=F) %>% 
  dplyr::mutate(percent= n/sum(n))
Enh.Prom.pairs.sig.summary.gene.prom <- Enh.Prom.pairs.sig.summary.gene.prom[Enh.Prom.pairs.sig.summary.gene.prom$count==1,]

Enh.Prom.pairs.sig.summary.gene.dist <- Enh.Prom.pairs.sig.summary[Enh.Prom.pairs.sig.summary$DistanceBin!="[0, 10000)",] %>% 
  dplyr::group_by(Gene, method) %>% 
  dplyr::count(count, .drop=F) %>% 
  dplyr::mutate(Percent= n/sum(n))
Enh.Prom.pairs.sig.summary.gene.dist <- Enh.Prom.pairs.sig.summary.gene.dist[Enh.Prom.pairs.sig.summary.gene.dist$count==1,]
```

Dotplot
```{r}
test <-colorRamp2(c(0.2,1), c( "#8b5b17" ,"black"))
test(0.8)
Gene.plot <- ggplot(Enh.Prom.pairs.sig.summary.gene.dist,aes(x=method,y=Gene)) +geom_point(aes(size=n,color=Percent))+
  scale_size(range = c(1,6))+
  TG_theme()+
  xlab("")+
  TG_theme(aspect.ratio=1)+
  scale_color_gradient(low="#d12027", high="black")+
  #scale_color_gradient(low="#b81965", high="black")+
  theme(axis.title.y=element_blank(), axis.title.x=element_blank(), axis.text.y = element_text(size=14), legend.text = element_text(size = 10),  legend.title = element_text(size=16, face = "bold"))

ggsave(path="output/plots/figure1", filename="K562_enhancer_annotation_gene_based.pdf", plot = Gene.plot, device = "pdf", dpi=300)

```

### 1.3.5 Example of situation where only interaction-based annotation identify elements 
Browser track of the Gata1 locus

### 1.3.6 False positivie analysis

Preparations
```{r}
dist_cutoff <-25000

#Select most signifcant
Enh.anno.ICEA.50 <- Enh.anno.ICEA[!(-log10(Enh.anno.ICEA$Interaction_score)<int.score.50 & Enh.anno.ICEA$Annotation_method=="Interaction_anno"),]
Enh.anno.ICEA.25 <- Enh.anno.ICEA[!(-log10(Enh.anno.ICEA$Interaction_score)<int.score.25 & Enh.anno.ICEA$Annotation_method=="Interaction_anno"),]
Enh.anno.ICEA.5 <- Enh.anno.ICEA[!(-log10(Enh.anno.ICEA$Interaction_score)<int.score.5 & Enh.anno.ICEA$Annotation_method=="Interaction_anno"),]

#Add stats for filtered
Enh.Prom.pairs.sig$ICEA_50 <- ifelse(Enh.Prom.pairs.sig$id %in% Enh.anno.ICEA.50$id, 1, 0)
Enh.Prom.pairs.sig$ICEA_25 <- ifelse(Enh.Prom.pairs.sig$id %in% Enh.anno.ICEA.25$id, 1, 0)
Enh.Prom.pairs.sig$ICEA_5 <- ifelse(Enh.Prom.pairs.sig$id %in% Enh.anno.ICEA.5$id, 1, 0)

#Group sig paris based on distance
Enh.Prom.pairs.sig.short <- Enh.Prom.pairs.sig[abs(Enh.Prom.pairs.sig$DistanceTSS)<=dist_cutoff,]
Enh.Prom.pairs.sig.long <- Enh.Prom.pairs.sig[abs(Enh.Prom.pairs.sig$DistanceTSS)>dist_cutoff,]


#Identified enhancers
Enh.Prom.pairs.sig.short.agg <- aggregate(.~Gene, Enh.Prom.pairs.sig.short[,c(1,29:36)], sum)
Enh.Prom.pairs.sig.short.agg.tot <- aggregate(.~Gene, Enh.Prom.pairs.sig.short[,c(1,8)], sum)
Enh.Prom.pairs.sig.short.agg$tot <- Enh.Prom.pairs.sig.short.agg.tot$Significant

Enh.Prom.pairs.sig.long.agg <- aggregate(.~Gene, Enh.Prom.pairs.sig.long[,c(1,29:36)], sum)
Enh.Prom.pairs.sig.long.agg.tot <- aggregate(.~Gene, Enh.Prom.pairs.sig.long[,c(1,8)], sum)
Enh.Prom.pairs.sig.long.agg$tot <- Enh.Prom.pairs.sig.long.agg.tot$Significant

Enh.prom.n <- rbind(data.frame(method=colnames(Enh.Prom.pairs.sig.short.agg[,2:9]), value=colMeans(Enh.Prom.pairs.sig.short.agg[,2:9]), group="proximal"), data.frame(method=colnames(Enh.Prom.pairs.sig.long.agg[,2:9]), value=colMeans(Enh.Prom.pairs.sig.long.agg[,2:9]), group="distal"))

Enh.prom.nf <- rbind(data.frame(method=colnames(Enh.Prom.pairs.sig.short.agg[,2:9]), value=colMeans(Enh.Prom.pairs.sig.short.agg[,2:9]/Enh.Prom.pairs.sig.short.agg$tot), group="proximal"), data.frame(method=colnames(Enh.Prom.pairs.sig.long.agg[,2:9]), value=colMeans(Enh.Prom.pairs.sig.long.agg[,2:9]/Enh.Prom.pairs.sig.long.agg$tot), group="distal"))


#False positives
Enh.anno.HOMER$sig <- ifelse(Enh.anno.HOMER$id %in% Enh.Prom.pairs.sig.f$id, 1, 0)
Enh.anno.GREAT$sig <- ifelse(Enh.anno.GREAT$id %in% Enh.Prom.pairs.sig.f$id, 1, 0)
Enh.anno.ICEA$sig <- ifelse(Enh.anno.ICEA$id %in% Enh.Prom.pairs.sig.f$id, 1, 0)
Enh.anno.overlap$sig <- ifelse(Enh.anno.overlap$id %in% Enh.Prom.pairs.sig.f$id, 1, 0)
Enh.anno.noprox$sig <- ifelse(Enh.anno.noprox$id %in% Enh.Prom.pairs.sig.f$id, 1, 0)
Enh.anno.ICEA.50$sig <- ifelse(Enh.anno.ICEA.50$id %in% Enh.Prom.pairs.sig.f$id, 1, 0)
Enh.anno.ICEA.25$sig <- ifelse(Enh.anno.ICEA.25$id %in% Enh.Prom.pairs.sig.f$id, 1, 0)
Enh.anno.ICEA.5$sig <- ifelse(Enh.anno.ICEA.5$id %in% Enh.Prom.pairs.sig.f$id, 1, 0)

Enh.anno.HOMER$group <- ifelse(abs(Enh.anno.HOMER$Distance.to.TSS)<=dist_cutoff, "short", "long")
Enh.anno.GREAT$group <- ifelse(abs(Enh.anno.GREAT$DistanceTSS)<=dist_cutoff, "short", "long")
Enh.anno.ICEA$group <- ifelse(abs(Enh.anno.ICEA$Distance_to_TSS)<=dist_cutoff, "short", "long")
Enh.anno.overlap$group <- ifelse(abs(Enh.anno.overlap$Distance_to_TSS)<=dist_cutoff, "short", "long")
Enh.anno.noprox$group <- ifelse(abs(Enh.anno.noprox$Distance_to_TSS)<=dist_cutoff, "short", "long")
Enh.anno.ICEA.50$group <- ifelse(abs(Enh.anno.ICEA.50$Distance_to_TSS)<=dist_cutoff, "short", "long")
Enh.anno.ICEA.25$group <- ifelse(abs(Enh.anno.ICEA.25$Distance_to_TSS)<=dist_cutoff, "short", "long")
Enh.anno.ICEA.5$group <- ifelse(abs(Enh.anno.ICEA.5$Distance_to_TSS)<=dist_cutoff, "short", "long")

#short
Enh.anno.HOMER.short  <- Enh.anno.HOMER[Enh.anno.HOMER$group=="short",] %>% group_by(Gene.Name) %>% summarize(count = sum(sig != 0)/n())
Enh.anno.GREAT.short  <- Enh.anno.GREAT[Enh.anno.GREAT$group=="short",] %>% group_by(Gene) %>% summarize(count = sum(sig != 0)/n())
Enh.anno.ICEA.short  <- Enh.anno.ICEA[Enh.anno.ICEA$group=="short",] %>% group_by(Gene) %>% summarize(count = sum(sig != 0)/n())
Enh.anno.overlap.short  <- Enh.anno.overlap[Enh.anno.overlap$group=="short",] %>% group_by(Gene) %>% summarize(count = sum(sig != 0)/n())
Enh.anno.noprox.short  <- Enh.anno.noprox[Enh.anno.noprox$group=="short",] %>% group_by(Gene) %>% summarize(count = sum(sig != 0)/n())
Enh.anno.ICEA.50.short  <- Enh.anno.ICEA.50[Enh.anno.ICEA.50$group=="short",] %>% group_by(Gene) %>% summarize(count = sum(sig != 0)/n())
Enh.anno.ICEA.25.short  <- Enh.anno.ICEA.25[Enh.anno.ICEA.25$group=="short",] %>% group_by(Gene) %>% summarize(count = sum(sig != 0)/n())
Enh.anno.ICEA.5.short  <- Enh.anno.ICEA.5[Enh.anno.ICEA.5$group=="short",] %>% group_by(Gene) %>% summarize(count = sum(sig != 0)/n())

Enh.anno.short <- merge(merge(merge(merge(merge(merge(merge(merge(data.frame(Gene=unique(Enh.Prom.pairs.sig.short$Gene)), Enh.anno.HOMER.short, by.x="Gene", by.y="Gene.Name", all.x=T), Enh.anno.GREAT.short, by="Gene", all.x=T),Enh.anno.ICEA.short, by="Gene", all.x=T), Enh.anno.overlap.short, by="Gene", all.x=T),Enh.anno.noprox.short, by="Gene", all.x=T), Enh.anno.ICEA.50.short, by="Gene", all.x=T), Enh.anno.ICEA.25.short, by="Gene", all.x=T), Enh.anno.ICEA.5.short, by="Gene", all.x=T)
rownames(Enh.anno.short) <-Enh.anno.short$Gene
Enh.anno.short <- Enh.anno.short[,2:9]
colnames(Enh.anno.short) <- c("HOMER", "GREAT", "ICEA", "Overlap", "NoProx","ICEA_50", "ICEA_25", "ICEA_5")
colMeans(Enh.anno.short, na.rm=T)

#long
Enh.anno.HOMER.long  <- Enh.anno.HOMER[Enh.anno.HOMER$group=="long",] %>% group_by(Gene.Name) %>% summarize(count = sum(sig != 0)/n())
Enh.anno.GREAT.long  <- Enh.anno.GREAT[Enh.anno.GREAT$group=="long",] %>% group_by(Gene) %>% summarize(count = sum(sig != 0)/n())
Enh.anno.ICEA.long  <- Enh.anno.ICEA[Enh.anno.ICEA$group=="long",] %>% group_by(Gene) %>% summarize(count = sum(sig != 0)/n())
Enh.anno.overlap.long  <- Enh.anno.overlap[Enh.anno.overlap$group=="long",] %>% group_by(Gene) %>% summarize(count = sum(sig != 0)/n())
Enh.anno.noprox.long  <- Enh.anno.noprox[Enh.anno.noprox$group=="long",] %>% group_by(Gene) %>% summarize(count = sum(sig != 0)/n())
Enh.anno.ICEA.50.long  <- Enh.anno.ICEA.50[Enh.anno.ICEA.50$group=="long",] %>% group_by(Gene) %>% summarize(count = sum(sig != 0)/n())
Enh.anno.ICEA.25.long  <- Enh.anno.ICEA.25[Enh.anno.ICEA.25$group=="long",] %>% group_by(Gene) %>% summarize(count = sum(sig != 0)/n())
Enh.anno.ICEA.5.long  <- Enh.anno.ICEA.5[Enh.anno.ICEA.5$group=="long",] %>% group_by(Gene) %>% summarize(count = sum(sig != 0)/n())

Enh.anno.long <- merge(merge(merge(merge(merge(merge(merge(merge(data.frame(Gene=unique(Enh.Prom.pairs.sig.long$Gene)), Enh.anno.HOMER.long, by.x="Gene", by.y="Gene.Name", all.x=T), Enh.anno.GREAT.long, by="Gene", all.x=T),Enh.anno.ICEA.long, by="Gene", all.x=T), Enh.anno.overlap.long, by="Gene", all.x=T), Enh.anno.noprox.long, by="Gene", all.x=T), Enh.anno.ICEA.50.long, by="Gene", all.x=T), Enh.anno.ICEA.25.long, by="Gene", all.x=T), Enh.anno.ICEA.5.long, by="Gene", all.x=T)
rownames(Enh.anno.long) <-Enh.anno.long$Gene
Enh.anno.long <- Enh.anno.long[,2:9]
colnames(Enh.anno.long) <- c("HOMER", "GREAT", "ICEA", "Overlap", "NoProx","ICEA_50", "ICEA_25", "ICEA_5")
colMeans(Enh.anno.long, na.rm=T)

Enh.prom.fp <- rbind(data.frame(method=colnames(Enh.anno.short), value=colMeans(Enh.anno.short, na.rm=T), group="proximal"), data.frame(method=colnames(Enh.anno.long), value=colMeans(Enh.anno.long, na.rm=T), group="distal"))

```

Visualization
```{r}
#Number of sig identified per gene
Enh.prom.n$group <- factor(Enh.prom.n$group, levels=c("proximal", "distal"))
Enh.prom.n$method <- factor(Enh.prom.n$method, levels=c("HOMER", "GREAT", "Overlap","NoProx","ICEA","ICEA_50", "ICEA_25", "ICEA_5"))

icea.sup.n <-ggplot(data=Enh.prom.n, aes(x=group, y=value, fill=group)) +
  geom_bar(stat="identity", position=position_dodge(width = 0.5), width=0.8, color="black")+
  facet_grid(. ~ method)+
  xlab("")+TG_theme()+
  scale_y_continuous(name="# Enhancers identified", expand = c(0,0))+
  scale_fill_manual(values = c("black", "#78A056"))
ggsave(path="output/plots/figure1", filename="icea_sup_n.pdf", plot = icea.sup.n, device = "pdf", dpi=300)


#Fraction of identified
Enh.prom.nf$group <- factor(Enh.prom.nf$group, levels=c("proximal", "distal"))
Enh.prom.nf$method <- factor(Enh.prom.nf$method, levels=c("HOMER", "GREAT", "Overlap","NoProx","ICEA","ICEA_50", "ICEA_25", "ICEA_5"))

icea.sup.nf <-ggplot(data=Enh.prom.nf, aes(x=group, y=value, fill=group)) +
  geom_bar(stat="identity", position=position_dodge(width = 0.5), width=0.8, color="black")+
  facet_grid(. ~ method)+
  xlab("")+TG_theme()+
  scale_y_continuous(name="% Enhancers identified", expand = c(0,0))+
  scale_fill_manual(values = c("black", "#78A056"))
ggsave(path="output/plots/figure1", filename="icea_sup_nf.pdf", plot = icea.sup.nf, device = "pdf", dpi=300)


#Fraction of sign element
Enh.prom.fp$group <- factor(Enh.prom.fp$group, levels=c("proximal", "distal"))
Enh.prom.fp$method <- factor(Enh.prom.fp$method, levels=c("HOMER", "GREAT", "Overlap","NoProx","ICEA","ICEA_50", "ICEA_25", "ICEA_5"))

icea.sup.fp <-ggplot(data=Enh.prom.fp, aes(x=group, y=value, fill=group)) +
  geom_bar(stat="identity", position=position_dodge(width = 0.5), width=0.8, color="black")+
  facet_grid(. ~ method)+
  xlab("")+TG_theme()+
  scale_y_continuous(name="% sig.  Enhancers", expand = c(0,0))+
  scale_fill_manual(values = c("black", "#78A056"))
ggsave(path="output/plots/figure1", filename="icea_sup_fp.pdf", plot = icea.sup.fp, device = "pdf", dpi=300)

```



### 1.3.7 Fraction of enhancers with H3K27ac
Overlap of Enhancers and H3K27ac peaks
```{bash engine.opts='-l'}
cd output/data/figure1/
bash /mnt/data/common/tobias/tg/ICE_A/ICEA_analysis/script/general/bedtools_intersect_wa.sh K562_Enhancers_filtered.bed ../../../data/chip_cnr/GSM733656_hg19_wgEncodeBroadHistoneK562H3k27acStdPk.broadPeak K562_Enhancers_filtered_H3K27ac_overlap.bed
```

Calculate percentage of enhancers with H3K27ac
```{r}
Enh.h3k27ac <- read.table("output/data/figure1/K562_Enhancers_filtered_H3K27ac_overlap.bed", header=F, sep="\t", quote = "")

Enh.Prom.pairs.bed$h3k27ac <- ifelse(Enh.Prom.pairs.bed$peakID %in% Enh.h3k27ac$V4, 1, 0)
Enh.Prom.pairs.bed$sig <- ifelse(Enh.Prom.pairs.bed$peakID %in% Enh.Prom.pairs[Enh.Prom.pairs$Significant,]$peakID, "Sig", "NoSig")
Enh.h3k27ac.stats <- table(Enh.Prom.pairs.bed$sig, Enh.Prom.pairs.bed$h3k27ac)

Enh.h3k27ac.stats <- rbind(data.frame(Group="Sig", Perc=Enh.h3k27ac.stats["Sig","1"]/sum(Enh.h3k27ac.stats["Sig",])), 
                           data.frame(Group="NoSig", Perc=Enh.h3k27ac.stats["NoSig","1"]/sum(Enh.h3k27ac.stats["NoSig",])))


plot.Enh.h3k27ac<-ggplot(Enh.h3k27ac.stats , aes(x=Group, y=Perc, fill=Group))+
  geom_bar(stat="identity", color="black", width = 0.7)+
  scale_y_continuous(breaks=c(0,0.25, 0.5, 0.75, 1), limits=c(0,1), expand=c(0,0), labels = scales::percent)+
  TG_theme()+
  scale_fill_manual(values = c("white", "black"))

ggsave(path="output/plots/figure1", filename="K562_enhancer_H3K27ac.pdf", plot = plot.Enh.h3k27ac, device = "pdf", dpi=300)

```

Fraction of ICE-A identified with H3K27me3
```{r}
sum(unique(Enh.Prom.pairs.sig.short[Enh.Prom.pairs.sig.short$ICEA==1,]$peakID) %in% Enh.Prom.pairs.bed[Enh.Prom.pairs.bed$h3k27ac==1,]$peakID)/length(unique(Enh.Prom.pairs.sig.short[Enh.Prom.pairs.sig.short$ICEA==1,]$peakID))

sum(unique(Enh.Prom.pairs.sig.long[Enh.Prom.pairs.sig.long$ICEA==1,]$peakID) %in% Enh.Prom.pairs.bed[Enh.Prom.pairs.bed$h3k27ac==1,]$peakID)/length(unique(Enh.Prom.pairs.sig.long[Enh.Prom.pairs.sig.long$ICEA==1,]$peakID))

sum(unique(Enh.Prom.pairs.sig[Enh.Prom.pairs.sig$ICEA==1,]$peakID) %in% Enh.Prom.pairs.bed[Enh.Prom.pairs.bed$h3k27ac==1,]$peakID)/length(unique(Enh.Prom.pairs.sig[Enh.Prom.pairs.sig$ICEA==1,]$peakID))

```



## 1.4 Benchmarking: Cancer enhancers

### 1.4.1 Custom promoters
```{r}
#Load custom promoter regions
hg38.promoters <- read.table("data/other/tss/GREATv4.genes.hg38.tsv", header=F, sep="\t", quote = "")
colnames(hg38.promoters) <- c("ensembl", "chr", "tss", "strand", "symbol")
hg38.promoters$start <-hg38.promoters$tss-2500
hg38.promoters$end <-hg38.promoters$tss+2500
hg38.promoters$strand <- ifelse(hg38.promoters$strand=="+", 0, 1)

hg38.promoters.f <- hg38.promoters[,c(5,2,6,7,4)]
write.table(hg38.promoters.f, "output/data/figure1/cancer_enhancers/hg38_tss_2500.txt", quote = F, sep = "\t", row.names = F, col.names = F)
```

### 1.4.2 Load data
```{r}
#Enahncers
enh.val <-  data.frame(read_xlsx(path = "data/other/cancer_enhancers/essential_enhancers.xlsx", sheet = 6, skip = 1))
enh.val$peakID <- paste0(enh.val$chr, ":", enh.val$start, "-", enh.val$end)
enh.val <-unique(enh.val[,c(2,3,4,11)])
write.table(enh.val, "output/data/figure1/cancer_enhancers/enh.bed", sep="\t", quote = F, col.names = F, row.names = F)

#Essential genes
genes <-  data.frame(read_xlsx(path = "data/other/cancer_enhancers/essential_genes.xlsx", sheet = 1))[,1:2]
colnames(genes) <- c("id", "type")
genes <- genes[genes$type=="POS_HCT_GECKO",]
genes$id <- sub("_.*", "", genes$id)
genes <- unique(genes)

#filter (within 2MB from enh)
genes.f <-hg38.promoters[hg38.promoters$symbol %in% genes$id,]
genes.f$reg_start <- genes.f$tss-2000000
genes.f$reg_end <- genes.f$tss+2000000
genes.f[genes.f$reg_start<0,]$reg_start <- 0
write.table(genes.f[,c(2,8,9)], "output/data/figure1/cancer_enhancers/gene_reg.bed", sep="\t", quote = F, col.names = F, row.names = F)

#Interactions
int.a549.1 <- read.table(file="data/other/cancer_enhancers/GSE161873_format_A549wtr1.10k.2.peaks.bedpe.txt", header = F, sep = "\t", fill = T)[,1:4]
colnames(int.a549.1) <- c("chr1", "s1", "e1", "loc2")
int.a549.1 <-  int.a549.1 %>% separate(loc2, c("chr2", "s2", "e2", "score"), ":|-|,", convert=T)

int.a549.2 <- read.table(file="data/other/cancer_enhancers/GSE161873_format_A549wtr2.10k.2.peaks.bedpe.txt", header = F, sep = "\t", fill = T)[,1:4]
colnames(int.a549.2) <- c("chr1", "s1", "e1", "loc2")
int.a549.2 <-  int.a549.2 %>% separate(loc2, c("chr2", "s2", "e2", "score"), ":|-|,", convert=T)

int.a549 <- rbind(int.a549.1, int.a549.2)
int.a549 <- int.a549[order(int.a549$score, decreasing = T),]
int.a549 <- int.a549[!duplicated(int.a549[,1:6]),]
write.table(int.a549, "output/data/figure1/cancer_enhancers/int_A549.bed", sep="\t", quote = F, col.names = T, row.names = F)

int.hct <- read.table(file="data/other/cancer_enhancers/GSE161873_format_HCT116wt.10k.2.peaks.bedpe.txt", header = F, sep = "\t", fill = T)[,1:4]
colnames(int.hct) <- c("chr1", "s1", "e1", "loc2")
int.hct <-  int.hct %>% separate(loc2, c("chr2", "s2", "e2", "score"), ":|-|,", convert=T)
int.hct <- int.hct[order(int.hct$score, decreasing = T),]
int.hct <- int.hct[1:nrow(int.a549),]
write.table(int.hct, "output/data/figure1/cancer_enhancers/int_HCT116.bed", sep="\t", quote = F, col.names = T, row.names = F)
```

### 1.4.3 Filer enhancers for close gene
```{bash engine.opts='-l'}
cd output/data/figure1/cancer_enhancers/
bash /mnt/data/common/tobias/tg/ICE_A/ICEA_analysis/script/general/bedtools_intersect_wa.sh enh.bed gene_reg.bed enh_f.bed
```

```{r}
enh_f_o <- read.table(file="output/data/figure1/cancer_enhancers/enh_f.bed", header = F, sep = "\t", fill = T)
enh_f_o <- unique(enh_f_o)
write.table(enh_f_o, "output/data/figure1/cancer_enhancers/enh_f.bed", sep="\t", quote = F, col.names = F, row.names = F)
```

### 1.4.4 Annotation

ICE-A annotation
```{bash engine.opts='-l'}
script_dir="script/figure1"
bash $script_dir/icea_cancer_enh.sh
```

```{r}
enh.anno.hct <- read.table(file="output/data/figure1/cancer_enhancers/enh_anno_HCT116/Peak_annotation/enh/enh_PLACseq_annotated.txt", header = T, sep = "\t", fill = T)
enh.anno.hct$peakID <- paste0(enh.anno.hct$Chr,":", enh.anno.hct$Start,"-",enh.anno.hct$End)
enh.anno.hct <- enh.anno.hct[,c(17, 9, 5, 10,13, 14)]
enh.anno.hct$id <- paste0(enh.anno.hct$peakID, "_", enh.anno.hct$Gene)
enh.anno.hct <- enh.anno.hct[!(enh.anno.hct$Peak_type=="Promoter" & enh.anno.hct$Annotation_method=="Interaction_anno"),]
enh.anno.hct <- enh.anno.hct[order(enh.anno.hct$Interaction_score),]
enh.anno.hct <- enh.anno.hct[!duplicated(enh.anno.hct$id),]

enh.anno.hct.int <- read.table(file="output/data/figure1/cancer_enhancers/enh_anno_HCT116_overlap/Peak_annotation/enh/enh_PLACseq_annotated.txt", header = T, sep = "\t", fill = T)
enh.anno.hct.int$peakID <- paste0(enh.anno.hct.int$Chr,":", enh.anno.hct.int$Start,"-",enh.anno.hct.int$End)
enh.anno.hct.int <- enh.anno.hct.int[enh.anno.hct.int$Annotation_method=="Interaction_anno",]
enh.anno.hct.int <- enh.anno.hct.int[,c(17, 9, 5, 10,13, 14)]
enh.anno.hct.int$id <- paste0(enh.anno.hct.int$peakID, "_", enh.anno.hct.int$Gene)
enh.anno.hct.int <- enh.anno.hct.int[order(enh.anno.hct.int$Interaction_score),]
enh.anno.hct.int <- enh.anno.hct.int[!duplicated(enh.anno.hct.int$id),]

enh.anno.a549 <- read.table(file="output/data/figure1/cancer_enhancers/enh_anno_A549/Peak_annotation/enh/enh_PLACseq_annotated.txt", header = T, sep = "\t", fill = T)
enh.anno.a549$peakID <- paste0(enh.anno.a549$Chr,":", enh.anno.a549$Start,"-",enh.anno.a549$End)
enh.anno.a549 <- enh.anno.a549[,c(17, 9, 5, 10,13, 14)]
enh.anno.a549$id <- paste0(enh.anno.a549$peakID, "_", enh.anno.a549$Gene)
enh.anno.a549 <- enh.anno.a549[!(enh.anno.a549$Peak_type=="Promoter" & enh.anno.a549$Annotation_method=="Interaction_anno"),]
enh.anno.a549 <- enh.anno.a549[order(enh.anno.a549$Interaction_score),]
enh.anno.a549 <- enh.anno.a549[!duplicated(enh.anno.a549$id),]
```

HOMER proximity annotation 
```{bash engine.opts='-l'}
cd output/data/figure1/cancer_enhancers

awk -v FS='\t' -v OFS='\t' '{print $2,"custom","exon",$3,$4,".",$5,".","transcript_id ""\x22"$1"\x22"}' hg38_tss_2500.txt > hg38_tss_2500.gtf

annotatePeaks.pl enh_f.bed hg38 -gtf hg38_tss_2500.gtf > enh_anno_prox.txt
awk -v FS='\t' -v OFS='\t' '{print $1, $16, $10}' enh_anno_prox.txt > enh_anno_prox_cut.txt
```

```{r}
enh.anno.p <- read.table("output/data/figure1/cancer_enhancers/enh_anno_prox_cut.txt", header=TRUE, sep="\t", quote="")
names(enh.anno.p)[1] <- 'peakID'
enh.anno.p <- enh.anno.p[!is.na(enh.anno.p),]
enh.anno.p$id <- paste0(enh.anno.p$peakID, "_", enh.anno.p$Gene)
enh.anno.p <- enh.anno.p[!duplicated(enh.anno.p$id),]
```

GREAT annotation
```{r}
enh.anno.great <- read.table("output/data/figure1/cancer_enhancers/enh_anno_great.txt", header=F, sep="\t", quote="")
colnames(enh.anno.great) <- c("peakID", "Gene")
enh.anno.great <- mutate(enh.anno.great,Gene = strsplit(Gene, ", ")) %>% unnest(Gene) %>% separate(Gene, c('Gene', 'DistanceTSS'), sep=" \\(")
enh.anno.great$DistanceTSS <- str_replace(enh.anno.great$DistanceTSS,
                        pattern = "\\)",
                        replacement = "")
enh.anno.great$DistanceTSS <- as.numeric(enh.anno.great$DistanceTSS)
enh.anno.great$id <- paste0(enh.anno.great$peakID, "_", enh.anno.great$Gene)
enh.anno.great <- data.frame(enh.anno.great[!duplicated(enh.anno.great$id),])
```


###  1.4.5 Stats
```{r}
#Prox
enh.prox <- enh.anno.p
enh.prox$ess <- ifelse(enh.prox$Gene %in% genes$id,1,0)

#great
enh.great <- enh.anno.great
enh.great$ess <- ifelse(enh.great$Gene %in% genes$id,1,0)

#icea - A548 - def
enh.anno.a549.def <- enh.anno.a549[!(enh.anno.a549$Peak_type=="Distal" & enh.anno.a549$Annotation_method=="Proximal_anno"), ]
enh.anno.a549.def$ess <- ifelse(enh.anno.a549.def$Gene %in% genes$id,1,0)

#icea - HCT - def
enh.anno.hct.def <- enh.anno.hct[!(enh.anno.hct$Peak_type=="Distal" & enh.anno.hct$Annotation_method=="Proximal_anno"), ]
enh.anno.hct.def$ess <- ifelse(enh.anno.hct.def$Gene %in% genes$id,1,0)

#icea - HCT only interactions
enh.anno.hct.intonly <- enh.anno.hct.int
enh.anno.hct.intonly$ess <- ifelse(enh.anno.hct.intonly$Gene %in% genes$id,1,0)


# %enh with ess gene
enh.ref <- length(unique(enh.prox[enh.prox$ess==1,]$peakID))/length(unique(enh_f_o$V4))*100
length(unique(enh.great[enh.great$ess==1,]$peakID))/length(unique(enh_f_o$V4))*100/enh.ref
length(unique(enh.anno.a549.def[enh.anno.a549.def$ess==1,]$peakID))/length(unique(enh_f_o$V4))*100/enh.ref
length(unique(enh.anno.hct.def[enh.anno.hct.def$ess==1,]$peakID))/length(unique(enh_f_o$V4))*100/enh.ref
length(unique(enh.anno.hct.intonly[enh.anno.hct.intonly$ess==1,]$peakID))/length(unique(enh_f_o$V4))*100/enh.ref

#Specificity: 
int.ref <- length(unique(enh.prox[enh.prox$ess==1,]$id))/length(unique(enh.prox$id))*100
length(unique(enh.great[enh.great$ess==1,]$id))/length(unique(enh.great$id))*100/int.ref
length(unique(enh.anno.a549.def[enh.anno.a549.def$ess==1,]$id))/length(unique(enh.anno.a549.def$id))*100/int.ref
length(unique(enh.anno.hct.def[enh.anno.hct.def$ess==1,]$id))/length(unique(enh.anno.hct.def$id))*100/int.ref
length(unique(enh.anno.hct.intonly[enh.anno.hct.intonly$ess==1,]$id))/length(unique(enh.anno.hct.intonly$id))*100/int.ref
```

